---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gsedread

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![License: CC BY-NC-ND 4.0](https://img.shields.io/badge/License-CC%20BY--NC--ND%204.0-lightgrey.svg)](https://creativecommons.org/licenses/by-nc-nd/4.0/)
<!-- badges: end -->

The goal of gsedread  is to read the Phase 1 and Phase 2 validation data of the project Global Scales for Early Development (GSED).

## Installation

Install the `gsedread` package from GitHub as follows:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("d-score/gsedread")
```

## How to use

You need access to the proper SharePoint site and sync the data to a local OneDrive. In the file `.Renviron` in your home directory add a line specifying the location of your synced OneDrive, e.g.,

```
ONEDRIVE_GSED='/Users/username/Library/CloudStorage/OneDrive-Sharedlibraries-...'
```

After saving `.Renviron`, (re)start R, and manually check whether you are able to read the OneDrive directory. The following R expression should return the contents of the OneDrive directory:

```{r}
head(dir(Sys.getenv("ONEDRIVE_GSED")))
```

Then set the proper paths **to** and **within** the OneDrive directory where the data are located. For example, for Phase 1 data, the paths are:

```{r}
onedrive <- Sys.getenv("ONEDRIVE_GSED")
path <- file.path("GSED Phase 1 Final Analysis",
                  "GSED Final Collated Phase 1 Data Files 18_05_22")
```

The high-level function `read_gsed_fixed()` reads and combines SF, LF and BSID data files. For the Phase 1 data, the function is called as follows:

```{r read_gsed_fixed}
library(gsedread)
phase1 <- read_gsed_fixed(onedrive, path, phase = 1)
```

The `read_gsed_fixed()` functions returns a list with two elements: `visits` and `responses`. 

The `visits` element is a data frame with the visit information, with the following structure: 

```{r}
str(phase1$visits)
```

There are `r nrow(phase1$visits)` visits in the data, with `r length(unique(phase1$visits$subjid))` unique GSED IDs.
A visit is defined by a combination of `subjid`, `agedays`, `vist_type` and `ins`.

The `responses` element is a data frame with the responses, with the following structure:

```{r}
str(phase1$responses)
```

There are `r nrow(phase1$responses)` responses in the data, with `r length(unique(phase1$responses$subjid))` unique GSED IDs.
A response is defined by a combination of `subjid`, `agedays`, `vist_type` and `item`. The column named `response` is the 0/1 score on the item.

## Lower-level functions

You can assess lower-level functions for reading data files, such as the Short Form (SF), by the `read_sf()` function. For example, to read the Phase 1 SF data files:

```{r}
sf <- read_sf(onedrive, path)
```

The resulting dataset is in a form much closer to the raw data. Normally, you would not need to use this function, as the `read_gsed_fixed()` function reads and processes the data for you, but it can help to diagnose issues with the data. Similar functions exist for the Long Form (LF) and BSID data files, `read_lf()` and `read_bsid()`, respectively.

## Operations

The package reads and processes GSED data. It does not store data. The `read_sf()`, `read_lf()` and `read_bsid()` functions take the following actions:

1. Construct the paths to the files OneDrive sync file
2. Read all specified datasets in a list
3. Internally specofy the desired format for each column
4. Specify the available date and data-time formats per file
5. Recode empty, `NA`, `-8888`, `-8,888.00` and `-9999` values as `NA`
6. Repair problems with mixed data-time formats in the adaptive Pakistan data
7. Stack the datasets to one tibble and adds columns `file` and `adm`
8. Remove records without a `GSED_ID`

In addition, the `read_gsed_fixed()` function performs the following actions:

9. Account for the different item orders between Phase 1 and Phase 2 for SF and LF by means of the `phase` argument
10. Add the `ins` field to `visits`, providing a short hand for instrument (`"sf"`, `"lf"`, `"bsid"`)
11. Transform the data for consistency, efficiency and clarity
12. Remove duplicates and sorts the data
13. Rename all item names to conform to the 9-position GSED convention
14. Split the data into `visits` and `responses` data frames, where `visits` contains the visit information and `responses` contains the item responses.
15. Remove `gpamoc008` (Clench fist)
16. Remove responses for `gtolgd002`, `gtolgd003`, `gtolgd004`, `gtolgd006`, `gtolgd007` and `gtolgd008` after the age of six months
17. Make sure that all responses are coded as 0 or 1.

Item renaming with `rename_variables()` relies on the item translation table at 
<https://github.com/D-score/gsedread/blob/main/inst/extdata/itemnames_translate.txt>.

## Additional cleaning and data repair

The `read_gsed_fixed()` reads the data per phase. After combining the data from the different phases, the functions `repair_visits()` and `repair_responses` should be run to ensure that the data are in a consistent format.

## Data governance

The Global Scales of Early Development (GSED) study is an international project coordinated by the World Health Organization (WHO). The data are stored in a secure SharePoint site and are shared within the GSED team for research purposes only. Team members are not allowed to share the data with third parties. See <https://www.who.int/publications/i/item/WHO-MSD-GSED-package-v1.0-2023.1> for information on the GSED study.

## Acknowledgement

This study was supported by the Bill & Melinda Gates Foundation. The contents are the sole responsibility of the authors and may not necessarily represent the official views of the Bill & Melinda Gates Foundation or other agencies that may have supported the primary data studies used in the present study. 

