---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gsedread

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of gsedread  is to read validation data of the project Global Scales for Early Development (GSED).

## Installation

<!-- If you have been marked as a collaborator on GitHub for this repository, generate a personal access token as in <https://github.com/settings/tokens>. Add a line  -->

<!-- ```{r eval=FALSE} -->
<!-- GITHUB_PAT=ghp_vC82..................... -->
<!-- ``` -->

<!-- with your token in the file `.Renviron` in your home directory. Restarting R adds the environmental variable GITHUB_PAT to your session. Then install the `gsedread` package from GitHub as follows:  -->

Install the `gsedread` package from GitHub as follows:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("d-score/gsedread")
```

There is no CRAN version.

## Example

You need access to the WHO SharePoint site and sync the data to a local OneDrive. In the file `.Renviron` in your home directory add a line specifying the location of your synced OneDrive, e.g.,

```
ONEDRIVE_GSED='/Users/username/Library/CloudStorage/OneDrive-Sharedlibraries-WorldHealthOrganization/CAVALLERA, Vanessa - GSED Validation 2021_phase I'
```

After setting the environmental variable `ONEDRIVE_GSED`, restart R, and manually check whether you are able to read the OneDrive directory.

```{r}
dir(Sys.getenv("ONEDRIVE_GSED"))
```

The following commands reads all SF data from `GSED Final Collated Phase 1 Data Files 18_05_22` directory and returns a tibble with one record per administration.

```{r read_sf}
library(gsedread)
data <- read_sf()
dim(data)
```

Count the number of records per file:

```{r}
table(data$file)
```

Process variable names user-friendly alternative:

```{r}
rename_vector(colnames(data)[c(1:3, 19, 21:25)], lexout = "gsed2", trim = "Ma_SF_")
```


## Operations

The package reads and processes GSED data. It does not store data. The `read_sf()` and `read_lf()` functions takes the following actions:

1. Constructs the paths to the files OneDrive sync file;
2. Reads all specified datasets in a list;
3. Internally specifies the desired format for each column;
4. Specifies the available date and data-time formats per file;
5. Recodes empty, `NA`, `-8888`, `-8,888.00` and `-9999` values as `NA`;
6. Repairs problems with mixed data-time formats in the adaptive Pakistan data;
7. Stacks the datasets to one tibble and adds columns `file` and `adm`;
8. Removes records without a `GSED_ID`.

Item renaming with `rename_variables()` relies on the item translation table at 
<https://github.com/D-score/gsedread/blob/main/inst/extdata/itemnames_translate.tsv>.

## Acknowledgement

This study was supported by the Bill & Melinda Gates Foundation. The contents are the sole responsibility of the authors and may not necessarily represent the official views of the Bill & Melinda Gates Foundation or other agencies that may have supported the primary data studies used in the present study. 
